### 2025-04-10 paper/header.tex
pushing updates preparing for final review

% **************GENERATED FILE, DO NOT EDIT**************

\title{BlankLocalizationCore.jl: implementing blank localization in Julia}

\author[1, 2]{Tamás Cserteg}
\author[1]{András Kovács}
\author[1]{József Váncza}
%\author[1, 3]{József Váncza}
\affil[1]{HUN-REN Institute for Computer Science and Control (SZTAKI), Budapest, Hungary}
\affil[2]{Doctoral School of Informatics, ELTE Eötvös Loránd University, Budapest, Hungary}
%\affil[3]{Department of Manufacturing Science and Technology, Budapest University of Technology and Economics, Budapest, Hungary}

\keywords{Julia, Optimization, Machining, Blank localization}

\hypersetup{
pdftitle = {BlankLocalizationCore.jl: implementing blank localization in Julia},
pdfsubject = {JuliaCon 2023 Proceedings},
pdfauthor = {Tamás Cserteg, András Kovács, József Váncza},
pdfkeywords = {Julia, Optimization, Machining, Blank localization},
}